---
title: "Multiple Linear Regression Project on Medical Cost Personal Datasets"
output:
  html_document:
    df_print: paged
---
## Multiple Linear Regression Project on Medical Cost Personal Datasets - Applied Linear Regression
Astrid Wang

December 18, 2020

datasource: https://www.kaggle.com/mirichoi0218/insurance or https://gist.github.com/meperezcuello/82a9f1c1c473d6585e750ad2e3c05a41 

The report paper including method discussion and reference is in a separate docs file
```{r}
#installed.packages("car")
#install.packages("olsrr")
#install.packages("reshape2")
#install.packages("tidyverse")
#install.packages("caret")
#install.packages("MuMIn")
#library(kenlab)
library(MuMIn)
library(reshape2)
library(ggplot2)
library("olsrr")
library(car)
library(caret)
df<- read.csv(file = 'insurance2.csv')
head(df)
```
```{r}
df <- transform(sex=factor(sex),smoker=factor(smoker),region=factor(region),df)
f1<-lm(charges~age+sex+bmi+children+smoker+region,data=df) # base model, or full model
summary(f1)
contrasts(df$sex)
contrasts(df$smoker)
contrasts(df$region)
```
```{r}
vif(f1)
plot(density(df$charges))
plot(density(log(df$charges)))
plot(f1)
```
```{r}
table(df$children) # try to understand the dataframe
plot(df$age,df$children)
plot(log(df$age),log(df$charges))
```

## Apply backward elimination for full model
```{r}
f2<-ols_step_backward_p(f1,details = TRUE) 

```

```{r}

f2<-lm(charges ~ age  + bmi + children + smoker+region, data = df)
summary(f2)
```
## Transformation Attemptation
```{r}
summary(powerTransform(f2)) # try power transformations but failed
invResPlot(f2)
```


```{r}
f6<-lm(log(charges) ~ age  + bmi + children + smoker+region, data = df) # use log instead of 1/3 power for the response variable because it makes more quantiles fashion the Normal distribution
summary(f6)
plot(f6)
```

## Add Weight

```{r}
df2<-suppressWarnings(transform(df, SD=apply(df,1, sd, na.rm = TRUE))) # new column sd
head(df2)
f3<-lm(log(charges) ~ age  + bmi + children + smoker+region, weights = 1/SD^2,data = df2)
summary(f3)
plot(f3) # increased normality a lot
```


```{r}
f7<-lm((charges)^(1/3) ~ age  + bmi + children + smoker+region, weights = 1/SD^2,data = df2) # normality issue did not solved by 1/3 power transformation, therefore log transformation is a good chioce. $
summary(f7)
plot(f7)
```


```{r}
pairs(~charges+age+bmi+ children ,data=df,main="Insurance Scatterplot Matrix")
f4<-lm(log(charges) ~ age  + bmi + children + smoker+region+age:children+bmi:smoker+age:bmi+age:smoker+bmi:children+children:smoker, weights = 1/SD^2,data = df2) # add all possible interaction terms
summary(f4)
plot(f4)
```
```{r}
ols_step_backward_p(f4,details = TRUE) # apply backward elimination again
```
```{r}
f5<-lm(log(charges) ~ age  + bmi + children + smoker+region+age:children+bmi:smoker+age:smoker+children:smoker, weights = 1/SD^2,data = df2)
summary(f5) # final model, the good one
```
```{r}
residualPlot(f5)
plot(f5)
```
## Questions and Findings
0. descriptive stat
```{r}
summary(df)
```

1. correlation matrix heatmap
```{r}
#head(df2)
df2$sex <- ifelse(df2$sex=="male", 1, 0)
df2$smoker <- ifelse(df2$smoker=="yes", 1, 0)
head(df2)
```

```{r}

df3 <- df2[, c(1,2,3,4,5,7)]
head(df3)
cormat <- round(cor(df3),2)
head(cormat)
melted_cormat <- melt(cormat)
head(melted_cormat)
ggplot(data = melted_cormat, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile()
```
```{r}
get_lower_tri<-function(cormat){
    cormat[upper.tri(cormat)] <- NA
    return(cormat)
}
get_upper_tri <- function(cormat){
    cormat[lower.tri(cormat)]<- NA
    return(cormat)
}
upper_tri <- get_upper_tri(cormat)
upper_tri
```
```{r}
melted_cormat <- melt(upper_tri, na.rm = TRUE)
ggplot(data = melted_cormat, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "red", high = "blue", mid = "pink", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Correlation Heatmap") +
  theme_minimal()+ 
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1))+
 coord_fixed()

```



2. box plot of smokers, regions
```{r}
boxplot(charges~smoker,
data=df,
main="Charges for Non-Smokers vs Smokers",
xlab="Smoking Habit",
ylab="Medical Insurance Cost",
col="orange",
border="brown"
)
```
```{r}
boxplot(charges~region,
data=df,
main="Charges for US Regions",
xlab="US Regions",
ylab="Medical Insurance Cost",
col="orange",
border="brown"
)
```


3. what is the model?
is there a positive linear relationship between log(charges) and age? CI? 
H0: B1=0, H1:B1>0. p-value < 2e-16 which is < 0.05 therefore reject H0, yes. 
```{r}
confint(f5,"age",level=0.95)
```
with smoking? 
```{r}
exp(1.5416300) # a smoker will spend $4.67 more than a non-smoker. 
```

with children?
```{r}
confint(f5,"children",level=0.95) # and yes
```
bmi?
```{r}
confint(f5,"bmi",level=0.95) #No, not in the final model
```


4. what expenses would you forecast, for a 22-year-old female non-smoker comes from northwest region with 0 children and 17.9 bmi? What is the 95% prediction interval?
```{r}
df4 <- data.frame(smoker = "no", children = 0, age = 22,bmi=17.9,region="northwest")
p1<-predict(f5, df4, interval = 'prediction')
p1
exp(p1)
```

## Model Selection table
not fitted in the same dataframe because I added a new column for the weight but the predictors did not change. 
```{r}
model.sel(f1, f2, f3, f4, f5, f6, f7, rank = AIC)
```

## ANOVA table for f5 which is the final model
```{r}
#parameterEstimates(f5)
anova(f5)
summary(f5)
```



